\paperTitle

### 2014-05-07 doc/icml2014/assumption-proof.tex
Included section about tensor projection and how to get Y in the assumption proof

@@

% \paragraph{Constructing $\mYpp{c}{\Pa(h_c)}$}
% 
% The algorithm constructs an expression for $\mYpp{t}{1}$ entirely as
%   tensor multiplications of terms of the form $\mYpp{c}{\Pa(h_c)}$ for
%   some hidden variable $h_c$. 
% The individual $\mYpp{c}{\Pa(h_c)}$ can be easily computed from the
%   clique marginals $Z_\sC$ as follows.
% 
% Without loss of generality, let $\sC \eqdef \{c\} \union \Pa(h_c)$ be a clique
%   in the graph $\sG$.
% If instead $\sC \subseteq \sC_1 \union \sC_2$ for some $\sC_1, \sC_2 \in
%   \sG$, we can simply construct $Z_{\sC}$ from the union of $\sC_1$ and
%   $\sC_2$,
% \begin{align*}
%   Z_\sC &= Z_{\sC_1} \times_{\del \sC} Z_{\sC_2},
% \end{align*}
% where $\del \sC = (\sC_1 \union \sC_2) \setminus \sC$.
% 
% Proceeding,
% \begin{align*}
% \mYpp{c}{\Pa(h_c)} 
%   &\eqdef \Pr(h_c \given \Pa(h_c)) \\
%   &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
%   &= Z_\sC \times_{\Pa(h_c)} \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1},
% \end{align*}
% where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
% tensor product with $I^{\otimes p}$:
% \begin{align*}
%   \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj].
% \end{align*}

>>>>

% Tour guide
Going forward, we will describe $\mOpp{v}{i}$ recursively in terms of
  tensor multiplications of conditional distributions $\Pr(h_c \given
  \Pa(h_c))$ in a variable elimination ordering.
We then show that the above assumption implies the unfoldings of the
  conditional distributions are full rank as well.
The analysis also tracks singular values, allowing us to present
  a sample complexity bound that highlights an exponential dependence on
  the size of the variable elimination set.

@@

\paragraph{Rank conditions for $\mYpp{c}{\Pa(h_c)}$}

\algorithmref{Y} constructs an expression for $\mYpp{t}{1}$ entirely as
  tensor multiplications of terms of the form $\mYpp{c}{\Pa(h_c)}$ for
  some hidden variable $h_c$. 
The singular values of unfoldings of individual $\mYpp{c}{\Pa(h_c)}$ can
  be easily computed from the clique marginals $Z_\sC$ as follows.

\begin{lemma} 
  Let $\sC_0 \eqdef {h_c} \union \Pa(h_c) \subseteq \sC_1 \union \cdots
  \union \sC_\ell$ for $\sC_1, \cdots, \sC_\ell \in \sG$. Note that
  $\sC_0$ need not belong in the graph.

  Then, for the unfolding $\{c\}$, we get 
  \begin{align*}
    \sigma_{\min}(\mYpp{c}{\Pa(c)}\munf{I}) &\ge \max_{i\in[\ell]}\{\sigma_{\min}(Z_{\sC_i}\munf{I_i})\},
  \end{align*}
  where $i$ is the subset of $I$ present in clique $\sC_i$.
\end{lemma} 
\begin{proof}
Without loss of generality, let $\sC \eqdef \{c\} \union \Pa(h_c)$ be a clique
  in the graph $\sG$.
If instead $\sC \subseteq \sC_1 \union \sC_2$ for some $\sC_1, \sC_2 \in
\sG$, then $Z_{\sC}$ is the marginalization of some $Z_{\sC_1 \union \sC_2}$
(which may not be a parameter in our model).

\lemmaref{tensor-projection} states that when a tensor, $Z_{\sC_1 \union
\sC_2}$, is projected by a non-zero vector $v$ (in this case, $\ones$),
then the singular values of the projection $Z_\sC$ are preserved along
any unfolding $I$ of $Z_\sC$:
\begin{align*}
  \sigma_{\min}(Z_{\sC}\munf{I}) &\ge \sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I})v_{\min} \\
  \sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I}) &\ge
  \sigma_{\min}(Z_{\sC}\munf{I})v_{\max}\inv.
\end{align*}

In this case, we get that $\sigma_{\min}(Z_{\sC_1 \union \sC_2}\munf{I})
= \sigma_{\min}(Z_{\sC}\munf{I})v_{\max}\inv$.

Furthermore, assuming that any unfolding of $Z_{\sC_1}$ is full rank, we
have that $Z_{\sC_1 \union \sC_2}$ is also full rank along any common
unfolding.

Proceeding,
\begin{align*}
\mYpp{c}{\Pa(h_c)} 
  &\eqdef \Pr(h_c \given \Pa(h_c)) \\
  &= \frac{\Pr(h_c, \Pa(h_c))}{\Pr(\Pa(h_c))} \\
  \mYpp{c}{\Pa(h_c)}\munf{I}
  &= Z_\sC\munf{I} \diag{Z_\sC(\ones, \cdot, \ldots, \cdot)}^{-1},
\end{align*}
where $\diag(\sP)$ is a ``diagonal'' tensor constructed by taking the
tensor product with the identity matrix,
\begin{align*}
  \diag(\sP)[\vi,\vj] &\eqdef \sP[\vi] \delta[\vi,\vj].
\end{align*}

We have from the previous argument that $Z_\sC\munf{I}$ has full column
rank $k$; furthermore, $Z_\sC(\ones, \cdot, \ldots, \cdot)$ is
guaranteed to have at least $k$ non-zero elements because its complete
marginals, $\mPi{i} \succ 0 ~\forall h_i \in \sC$ have full support. 
Consequently, $\diag(Z_\sC(\ones, \cdot, \ldots, \cdot))$ has column rank $k$.
\todo{What happens if $\Pr(h_2, h_3) = 0$ at some point?}
\end{proof}

### 2014-05-07 doc/icml2014/tensor-multiplication.tex
Included section about tensor projection and how to get Y in the assumption proof

%In this section, we present basic results regarding two tensor
%operations we define, tensor multiplication, which is analogous to
%matrix multiplication, and the diagonal tensor.
In this section, we present basic results regarding tensor
multiplication, an operation analogous to matrix multiplication.
First, let us define some notation.

In the following, let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be
  a $m$-th order tensor and let $\sQ \in \Re^{d'_1 \times \ldots \times
  d'_n}$ be an $n$-th order tensor. %, where wlog, $m \ge n$. 
We use the notation $\Domain(\sP) = [d_1] \times \cdots \times [d_m]$
  to denote the domain of the indices of $\sP$.
Let $M$ be an index matching set, i.e. $M = \{ (a_1; b_1),
  (a_2, b_2), \ldots, (a_\ell, b_\ell) \}$, such that the corresopnding
  dimensions agree, i.e. $d_{a_i} = d'_{b_i}$ for all $i$. Let $D_{\ba}
  = [d_{a_1}] \times \cdots \times [d_{a_\ell}]$ and $D'_{\bb}
  = [d'_{b_1}] \times \cdots \times [d'_{b_\ell}]$ be the domains of the
  corresponding indices.

Finally, for some $\vi \in [d_1] \times \cdots \times [d_\ell]$ and $\vj
\in [d_{\ell + 1}] \times \cdots \times [d_m]$ and  we use the notation
$\sP[\vi \cdot \vj]_\pi \eqdef \sP[\pi(i_1, \cdot, i_\ell, j_{\ell+1},
\cdots j_m)]$ to conjoin indices $\vi$ and $\vj$ up to the permutation
$\pi$. 
%For our purposes, the specific ordering of the tensor's indices
%are not important and are usually referred to by variable naming.

@@

\begin{definition}[Tensor Multiplication]
  We define the tensor multiplication operation $\sR \eqdef P \times_{M}
  Q$ to be the $(m - \ell + n - \ell)$-th order tensor $\sR$ formed by
  conjoining $\sP$ and $\sQ$ followed by summing out the indices in $M$:
\begin{align*}
  \sR[\vi\cdot\vj] &\eqdef \sum_{\mathclap{\bu, \bv \in D_\ba \times D'_\bb}} \sP[\vi \cdot \bu]_{\pi_P} \sQ[\vj \cdot \bv]_{\pi_Q} ~ \delta_{u_1 v_1} \delta_{u_2 v_2} \ldots \delta_{u_\ell v_\ell},
\end{align*}
  for every $\vi \in \Domain(\sP) \setminus D_a$ and $\vj \in
  \Domain(\sQ) \setminus D'_b$, where $\pi_P$ and $\pi_Q$ are
  permutations of the indices that respect the matching $M$. For
  notational convenience, we will omit $\pi_P$, $\pi_Q$ in the sequel.
\end{definition}

@@

\paragraph{Unfoldings}
For a tensor $\sP$, as defined above, 
  given an index set $I \subset [m]$ we
  define an unfolding of a tensor $\sP$ along $I$,
 $\sP\munf{I}$, to be a matrix of dimensions $(\prod_{i \in I} d_i)
 \times (\prod_{j \not\in I} d_j)$. $\sP\munf{I}$ has elements
 $\sP\munf{I}[\vi;\vj] = \sP[\vi\cdot\vj]_{\pi}$, where $\pi$ is an
 appropriate permutation to align the indices of $\vi$ and $\vj$ with
 the ordering of $I$. 
For example, if $\sP$ were a $4$-th order tensor and $I = \{1,3\}$,
  $\sP\munf{13}[i_1,i_2; j_1,j_2] = \sP[i_1,j_1,i_2,j_2]$.
For notational convenience, we will omit the permutation $\pi$ subsequently.
Let $D_I = \bigtimes_{i \in I} [d_i]$ and  $\bar I \eqdef [m] \setminus
  I$ to represent the complement of indices. Note that $\sP\munf{I}
  = (\sP\munf{\bar I})^\top$.

The main result of this section is, similar to matrix multiplication, if
$\sR = \sP \times_{M} \sQ$, then the singular values of unfoldings
of $\sR$ are bounded by those of $\sP$ and $\sQ$.

@@

\begin{proof}
  Without loss of generality, let $M = \{(1,1), \ldots, (\ell,\ell)\}$,
  and $D_\ell = [d_1] \times \cdots \times [d_\ell]$. 
  Then, 
  \begin{align*}
    \sR[\vi\cdot\vj] &\eqdef \sum_{\vk \in D_\ell} \sP[\vk\cdot\vi_P\cdot\vj_P] \sQ[\vk\cdot\vi_Q\cdot\vj_Q],
  \end{align*}
  where $\vi_P, \vi_Q$, are the components of $\vi$ in $\sP$ and $\sQ$
  respectively and $\vj_P, \vj_Q$ is similarly defined.

Consider an $(I)$-unfolding of $\sR$,
  \begin{align*}
    \sR\munf{I}[\vi;\vj] 
       &= \sum_{\vk \in D_\ell} \sP[\vk\cdot \vi_P \cdot \vj_P] \sQ[\vk\cdot \vi_Q \cdot \vj_Q] \\
       &= \sum_{\vk \in D_\ell} \sP\munf{I_P}[\vi_P; \vj_P \cdot \vk] \sQ\munf{\bar I_Q}[\vj_Q ; \vi_Q \cdot \vk].
  \end{align*}
Note that the indices that were summed out belong in the complement of
$I_P$ and $I_Q$ respectively.

@@

Now, $\sP'$ is a $(m + 2 |I_Q|)$-th order tensor and similarly $\sQ'$
  is a $(n + 2 |\bar I_P|)$-th order tensor. 
For every identity matrix we tensor $\sP$ and $\sQ$ by, we include one of
  its indices in the new set $I'_P$ and $I'_Q$ respectively.
Note that ${\sP'}\munf{I'_P}$ is a $(I_P + I_Q)\times (|\bar I_P| + \ell
  + |I_Q|)$ matrix, and ${\sQ'}\munf{\bar I'_Q}$ is $(|\bar I_Q| + |\bar
  I_P|) \times (|\bar I_Q| + \ell + |\bar I_P|)$ matrix.
Then,
\begin{align*}
    \sR\munf{I}[\vi;\vj] 
    &= \sum_{\be_P, \be_Q, \vk}
    {\sP'}\munf{I'_P}[\vi_P \cdot \be'_Q; \vj_P \cdot \vk \cdot \be_Q ]  \\ 
    &\quad \hphantom{\sum_{\be_P, \be_Q, \vk}}
    {\sQ'}\munf{I'_Q}[\vi_Q \cdot \be'_P; \vj_Q \cdot \vk \cdot \be_P ] \\
    \sR\munf{I} &= {\sP'}\munf{I'_P} ({\sQ'}\munf{\bar I'_Q})^\top \\
                &= {\sP'}\munf{I'_P} {\sQ'}\munf{I'_Q}.
\end{align*}

@@

>>>>

%In this section, we present basic results regarding two tensor
%operations we define, tensor multiplication, which is analogous to
%matrix multiplication, and the diagonal tensor.
In this section, we present some results regarding a few tensor
operations, most significant of which is tensor multiplication, an
operation analogous to matrix multiplication.  
First, let us define some notation.

\paragraph{Notation}
In the following, let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be
  a $m$-th order tensor and let $\sQ \in \Re^{d'_1 \times \ldots \times
  d'_n}$ be an $n$-th order tensor. %, where wlog, $m \ge n$. 
We use the notation $\Domain(\sP) = [d_1] \times \cdots \times [d_m]$
  to denote the domain of the indices of $\sP$.
Let $M$ be an index matching set, i.e. $M = \{ (a_1; b_1),
  (a_2, b_2), \ldots, (a_\ell, b_\ell) \}$, such that the corresopnding
  dimensions agree, i.e. $d_{a_i} = d'_{b_i}$ for all $i$. Let $D_{\ba}
  = [d_{a_1}] \times \cdots \times [d_{a_\ell}]$ and $D'_{\bb}
  = [d'_{b_1}] \times \cdots \times [d'_{b_\ell}]$ be the domains of the
  corresponding indices.

Finally, for some $\vi \in [d_1] \times \cdots \times [d_\ell]$ and $\vj
\in [d_{\ell + 1}] \times \cdots \times [d_m]$ and  we use the notation
$\sP[\vi \ \vj]_\pi \eqdef \sP[\pi(i_1, \cdots, i_\ell, j_{\ell+1},
\cdots j_m)]$ to conjoin indices $\vi$ and $\vj$ up to the permutation
$\pi$.

\paragraph{Unfoldings}
For a tensor $\sP$, as defined above, 
  given an index set $I \subset [m]$ we
  define an unfolding of a tensor $\sP$ along $I$,
 $\sP\munf{I}$, to be a matrix of dimensions $(\prod_{i \in I} d_i)
 \times (\prod_{j \not\in I} d_j)$. $\sP\munf{I}$ has elements
 $\sP\munf{I}[\vi;\vj] = \sP[\vi\ \vj]_{\pi}$, where $\pi$ is an
 appropriate permutation to align the indices of $\vi$ and $\vj$ with
 the ordering of $I$. 
For example, if $\sP$ were a $4$-th order tensor and $I = \{1,3\}$,
  $\sP\munf{13}[i_1,i_2; j_1,j_2] = \sP[i_1,j_1,i_2,j_2]$.
For notational convenience, we will omit the permutation $\pi$ subsequently.
Let $D_I = \bigtimes_{i \in I} [d_i]$ and  $\bar I \eqdef [m] \setminus
  I$ represent the complement of indices. Note that $\sP\munf{I}
  = (\sP\munf{\bar I})^\top$.

%For our purposes, the specific ordering of the tensor's indices
%are not important and are usually referred to by variable naming.

@@

\paragraph{Tensor Multiplication}
First, let us formally define the tensor multiplication operation.

\begin{definition}[Tensor Multiplication]
  We define the tensor multiplication operation $\sR \eqdef P \times_{M}
  Q$ to be the $(m - \ell + n - \ell)$-th order tensor $\sR$ formed by
  conjoining $\sP$ and $\sQ$ followed by summing out the indices in $M$:
\begin{align*}
  \sR[\vi\ \vj] &\eqdef \sum_{\mathclap{\bu, \bv \in D_\ba \times D'_\bb}} \sP[\vi \  \bu]_{\pi_P} \sQ[\vj \  \bv]_{\pi_Q} ~ \delta_{u_1 v_1} \delta_{u_2 v_2} \ldots \delta_{u_\ell v_\ell},
\end{align*}
  for every $\vi \in \Domain(\sP) \setminus D_a$ and $\vj \in
  \Domain(\sQ) \setminus D'_b$, where $\pi_P$ and $\pi_Q$ are
  permutations of the indices that respect the matching $M$. For
  notational convenience, we will omit $\pi_P$, $\pi_Q$ in the sequel.
\end{definition}

@@

The main result of this section is that, similar to matrix
multiplication, if $\sR = \sP \times_{M} \sQ$, then the singular values
of unfoldings of $\sR$ are bounded by those of $\sP$ and $\sQ$.

@@

\begin{proof}
  Without loss of generality, let $M = \{(1,1), \ldots, (\ell,\ell)\}$,
  and $D_\ell = [d_1] \times \cdots \times [d_\ell]$. 
  Then, 
  \begin{align*}
    \sR[\vi\ \vj] &\eqdef \sum_{\vk \in D_\ell} \sP[\vk\ \vi_P\ \vj_P] \sQ[\vk\ \vi_Q\ \vj_Q],
  \end{align*}
  where $\vi_P, \vi_Q$, are the components of $\vi$ in $\sP$ and $\sQ$
  respectively and $\vj_P, \vj_Q$ is similarly defined.

Consider an $(I)$-unfolding of $\sR$,
  \begin{align*}
    \sR\munf{I}[\vi;\vj] 
       &= \sum_{\vk \in D_\ell} \sP[\vk\  \vi_P \  \vj_P] \sQ[\vk\  \vi_Q \  \vj_Q] \\
       &= \sum_{\vk \in D_\ell} \sP\munf{I_P}[\vi_P; \vj_P \  \vk] \sQ\munf{\bar I_Q}[\vj_Q ; \vi_Q \  \vk].
  \end{align*}
Note that the indices that were summed out belong in the complement of
$I_P$ and $I_Q$ respectively.

@@

Now, $\sP'$ is a $(m + 2 |I_Q|)$-th order tensor and similarly $\sQ'$
  is a $(n + 2 |\bar I_P|)$-th order tensor. 
For every identity matrix we tensor $\sP$ and $\sQ$ by, we include one of
  its indices in the new set $I'_P$ and $I'_Q$ respectively.
Note that ${\sP'}\munf{I'_P}$ is a $(I_P + I_Q)\times (|\bar I_P| + \ell
  + |I_Q|)$ matrix, and ${\sQ'}\munf{\bar I'_Q}$ is $(|\bar I_Q| + |\bar
  I_P|) \times (|\bar I_Q| + \ell + |\bar I_P|)$ matrix.
Then,
\begin{align*}
    \sR\munf{I}[\vi;\vj] 
    &= \sum_{\vj'_P, \vi'_Q, \vk}
    {\sP'}\munf{I'_P}[\vi_P \  \vi_Q; \vj'_P \  \vk \  \vi'_Q ]  \\ 
    &\quad \hphantom{\sum_{\be_P, \be_Q, \vk}}
    {\sQ'}\munf{I'_Q}[\vj_Q \  \vj_P; \vi'_Q \  \vk \  \vj'_P ] \\
    \sR\munf{I} &= {\sP'}\munf{I'_P} ({\sQ'}\munf{\bar I'_Q})^\top \\
                &= {\sP'}\munf{I'_P} {\sQ'}\munf{I'_Q}.
\end{align*}

@@

\paragraph{Tensor Projections}

We will now study the singular values of a tensor projection. 
Let $\sP \in \Re^{d_1 \times \ldots \times d_m}$ be a $m$-th order
  tensor. 
For some vector $v \succ 0$, let $\sQ$ be the $m-1$-th order tensor
  formed by projecting one of the axes of $\sP$ by $v$. 
Without loss of generality, let the projected axes be the last one, $\sQ
  \eqdef \sP(\cdot, \cdots, \cdot, v)$.
Then, we have the following lemma linking the singular values of $\sQ$
  and $\sP$.
\begin{lemma}[Tensor projection preserves rank]
  \label{lem:tensor-projection}
  Let $\sP, \sQ$ and $v$ be defined as above. For any unfolding $I
  \not\contains v$, 
\begin{align*}
  \sigma_{\max}(\sQ\munf{I}) &\le \sigma_{\max}(\sP\munf{I}) v_{\max} \\
  \sigma_{\min}(\sQ\munf{I}) &\ge \sigma_{\min}(\sP\munf{I}) v_{\min}.
\end{align*}
In particular, $\rank(\sP\munf{I}) = \rank(\sQ\munf{I})$.
\end{lemma}
\begin{proof}
  We will show that $\sQ\munf{I}$ is $\sP\munf{I}$ related by
  a projection operator. When $v \succ 0$, this projection operator has
  full column rank, entailing the result.

  By definition of $\sQ$, for any $\vx$
  \begin{align*}
    \sQ[\vi] &= \sum_{k=1}^{d_m} \sP[\vi, k] v[k].
  \end{align*}
  Consider the unfolding along $I$,
  \begin{align*}
    \sQ\munf{I}[\vi;\vj] &= \sum_{k=1}^{d_m} \sP\munf{I}[\vi;\vj\ k] v[k].
  \end{align*}

  We can represent this as matrix multiplication by taking the tensor
  product of $v$ with the $D_{\bar I \setminus \{m\}} \times D_{\bar
  I \setminus \{m\}}$ identity matrix, $\bI$. Note that $\bI \otimes v$ is a $
  D_{\bar I \setminus \{m\}} d_m \times D_{\bar I \setminus \{m\}}
  = D_{\bar I } \times D_{\bar I \setminus \{m\}}$ matrix.
  \begin{align*}
    \sQ\munf{I}[\vi;\vj] 
      &= \sum_{j'=\ones, k=1}^{\bd, d_m} \sP\munf{I}[\vi;\vj'\ k] (v[k] \delta[\vj';\vj]) \\
    \sQ\munf{I}
      &= \sP\munf{I} \bI \otimes v.
  \end{align*}
  If $v$ has no non-zero elements, $\bI \otimes v$ has full-column
  rank, with singular values $v_1, \cdots, v_{d_m}$.
\end{proof}
